Constraining the Starspot Filling Factor of Magnetically Active M67 Sub-Subgiant S1063

### 2021-05-18 document/ms.tex
:memo: Lots of content change to the discussion

Starspots have become a rich observational proxy for gauging the evolutionary state of stars that may be magnetically inflated without needing to estimate magnetic field strengths.  A full accounting of starspots requires careful consideration of geometrical degeneracies.  Monochromatic light curves can indicate spotted stars \citep{2014ApJS..211...24M} but only provide a lower limit on the differential spot coverage between the most and least spotted hemispheres. Longitudinally symmetric starspot geometries evade detection in monochromatic lightcurves alone \citep{2019AJ....157...64L}. Disk-integrated covering fractions can be obtained from TiO band observations \citep{oneal96,fang2016,2019AJ....158..101M}, but measuring the spot configuration typically requires Doppler imaging or interferometry studies restricted to only the brightest and most nearby sources \citep{roettenbacher16}.  Testing the next era of stellar activity models requires precision methodologies to measure spot-covering fractions of a larger number of targets, including SSGs.

Starspots emit a spectrum of their own at a lower temperature and with distinct absorption features compared to the ambient photosphere. The observed spectrum is therefore a composite of the spot and photosphere spectra, which can be deconvolved.  \citet{gullysantiago17} extend the \texttt{Starfish} \citep{czekala15} spectral inference framework to support composite spectra and two-temperature probabilistic spectral decomposition. This deconvolution process constrains the spot-covering fraction as well as the spot and photosphere temperatures. The resulting disk-average spot properties anchor a simultaneous light curve flux to the starspot coverage fraction. This anchored value serves as a benchmark for interpreting long-term (multi-year) flux variability \citep{neff95}. This strategy only requires high-resolution near-IR echelle spectroscopy and photometric monitoring, making it less demanding on telescope resources than either Doppler Imaging or interferometry. The technique is amenable to sources with narrow spectral lines ($v\sin{i}<10$ km/s) and stars with any stellar inclination, a distinct advantage over Doppler imaging.  Altogether these advantages dramatically increase the number of sources for which we can observationally constrain spot-covering fractions.

In this paper we demonstrate the power of this methodology by focusing on a single SSG system, S1063 \citep{1977A&AS...27...89S} in the open cluster M67. This system is a prototypical SSG, with a single-lined spectroscopic orbital period of 18.4 days \citep{geller17}, an X-ray luminosity of $1.3\times10^{31}$ erg s$^{-1}$ \citep{vandenberg99}, and a variable light curve in \textit{K2} and ASAS-SN (Section~\ref{sec:K2lightcurve}). \emph{Gaia} DR2 astrometry \citep{2016A&A...595A...1G, 2018A&A...616A...1G} indicates a parallax ($1.17\pm0.025 \;$mas) and proper motion for S1063 (\emph{Gaia} DR2 604921030968952832) consistent with other M67 members  approaching 100\% membership probability \citep{2018ApJ...869....9G}. We present our observational data products in Section~\ref{sec:observations} and subsequent analysis in Section~\ref{sec:analysis}. Our results are given in Section~\ref{sec:results} with Discussion in Section~\ref{sec:discussion} including additional discussion of the possible systematic uncertainties that may bias our results. Finally, our conclusions are outlined in Section~\ref{sec:conclusions}.

@@

To account for sensitivity changes, we calibrate the system-integrated throughput for \emph{Kepler} detector channels including the M67 field for Campaigns 5, 16, and 18. We measure aperture photometry for approximately 2000 isolated reference stars from the Full Frame Images, keeping only those stars that were observed in all three campaigns. Compared to Campaign 5, the reference stars have a median flux of $93.9\pm4.2\%$ in Campaign 16 and a median flux of $98.2\pm2.8\%$ in Campaign 18. Campaigns 5 and 18 were observed on the same detector channel while a different detector channel was used for Campaign 16, therefore the C16 offset is not a significant measure of detector degradation across campaigns. We use a Lomb-Scargle periodogram on each campaign independently and recover the same dominant period of 23.5 days in all three campaigns. The phase-folded light curves and periodograms are shown in Figure~\ref{fig:periodogram}.

\begin{figure*}[ht]
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=3in]{figures/AllCampaigns_Periodogram.pdf}   & \includegraphics[width=3in]{figures/AllCampaigns_Phased_Lightcurve.pdf}  \\
    \end{tabular}
    \caption{On the right, a Lomb-Scargle Periodogram of the light curve for each \emph{K2} campaign and on the left, the corresponding phased \emph{K2} light curves. The average period associated with maximum power across all three campaigns is P = 23.5 days.  }
    \label{fig:periodogram}
\end{figure*}

\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.95\textwidth]{figures/2020_K2_ASASSN_lcurve_2panel.pdf}
\caption{Four year lightcurve for S1063, normalized to a maximum flux of 1.00 as denoted by the gray dashed line. The mean flux of 0.895 is shown with the gray dotted line.  \emph{K2} Campaigns 5, 16, and 18 (densely-sampled red points) and ASAS-SN $V-$band (coarsely-sampled gray points) show approximately $5-17\%$ peak-to-valley photometric variations indicative of secularly evolving surface coverage of starspots.  The orange line shows a damped, driven harmonic oscillator \texttt{celerit\`e} model \citep{2017AJ....154..220F} trained on the \emph{K2} and ASAS-SN data, and then applied to the noisy ASAS-SN points. A standard-error uncertainty band is shown in shaded orange. The vertical blue bar in mid-2015 indicates the epoch of IGRINS data acquisition, which shows an approximate $8\%$ flux deficit compared to the global maximum flux in early 2017.}
\label{fig:lightcurve}
\end{figure*}

@@

%INTRO TEXT:
%Starspots emit a spectrum of their own at a lower temperature and with distinct absorption features compared to the ambient photosphere. The observed spectrum is a composite of the spot and photosphere spectra and can therefore be deconvolved. This deconvolution constrains the spot-covering fraction as well as the spot and photosphere temperatures. This is possible by performing a two-temperature probabilistic spectral decomposition using the spectral inference framework \texttt{Starfish} \citep{czekala15}, recently extended to support composite spectra \citep{gullysantiago17}. The resulting spot characterization can be used to anchor light curve information, providing insight into the overall stellar activity level and constraining the long-term spot coverage \citep{neff95}. This strategy only requires high-resolution echelle spectroscopy and photometric monitoring, dramatically increasing the number of sources for which we can observationally constrain spot-covering fractions.

%Constraining starspot covering fraction for stars such as S1063 requires a careful combination of light curve interpretation and spectral analysis using spectral decomposition tools such as \texttt{Starfish} \citep{czekala15} to benchmark epochs of starspot coverage in long-term photometric monitoring.
Characterizing the long-term starspot coverage of S1063 requires connecting coarse ground-based and precision space-based photometry---ASAS-SN and \emph{K2}---alongside precision IGRINS spectroscopy.  Here we lay out the calibration of these datasets to accommodate their wide range in data quality.

@@

For a first-order interpretation of light curve amplitudes, we posit that the lightcurve is starspot-dominated as opposed to facula-dominated.  Under this assumption, minimum light corresponds to the largest blockage of ambient flux and therefore the largest starspot coverage on the instantaneous hemisphere projected towards the observer \cite{basri18}.  Maximum light corresponds to the moments with the least---\emph{but not necessarily zero}---starspot coverage.

@@

This 13\% spot coverage is a minimum value based on an assumption that the maximum lightcurve flux results from an un-spotted star. Our spectral inference results indicate the spot coverage at the time of the IGRINS observation was, in fact, 32\%. This suggests that at maximum light over this 4-year period the spot coverage of S1063 was approximately 20\%, and not zero as a first-order interpretation of the light curve would imply.

The presence of starspots at the light curve maximum acts as a starspot baseline for interpreting the lightcurve. Over the four years shown in Figure~\ref{fig:lightcurve}, the average flux variation is approximately 5\%, corresponding to an average spot covering fraction variation from each observed hemisphere of $\pm$8\%. The lightcurve exhibits a global flux minimum 10\% lower than the flux at the time of the IGRINS observation, requiring a spot filling factor close to 45\% on the projected hemisphere at that time.

We therefore determine that over the time period from 2014--2018 S1063 has possessed a range of 20--45\% coverage fraction of spots, with an average close to 30\%. The IGRINS observation occurring near the mean global flux level fortuitously allows a robust characterization of the average state of S1063. These numbers possess formal uncertainties in the few percent range, and additional uncertainties from our reliance on spectral models.  But broadly speaking, our result is firm under our assumptions --- S1063 has a large persistent starspot population that ebbs and flows in its longitudinal symmetry, resulting in variation in the peak-to-valley modulation, but always possesses an approximately 30\% spot covering fraction.

@@

We believe that some of the discrepancy between our results and those predicted by the SPOTS models may be due to limitations of the SPOTS models as diagnostic tools for single sources. However, observational biases may also impact the filling factors determined through spectroscopic and photometric methods. The spectral decomposition technique used here is sensitive to larger spot filling factors \citep{gullysantiago17}, as long as the spots are of sufficient brightness to provide a significant spectral signature. We do detect a significant spot spectral signature in S1063, but we do not know if we are detecting the spectrum of the umbra or prenumbra. It is most likely that the penumbra is larger in area and warmer than the umbra for most stars \citep{1981ApJ...250..327V}. Spectral decomposition techniques are more sensitive to the warmer spot components and may therefore be constraining the prenumbra filling factor, which would be larger than the umbra filling factor. However, we do not know if the spot morphology of SSGs follows this assumption. If instead SSGs have a smaller ratio of prenumbra to umbra this spectral technique would under-sample the total spot coverage on the stellar surface.

Additionally, possible wavelength dependencies of spot size would lead to discrepant spot filling factors across photometric and spectroscopic determinations at different wavelengths. We note that photometry collected at different epochs can also impact the inferred photometric filling factor. Given our light curve analysis of S1063 we expect epoch-based filling factor uncertainties could be as large as $\sim$20\%, which could bring the two-temperature SED fit from \citet{leiner17} and our results into agreement.

A summary of the surface and stellar parameters for S1063 are given in Table~\ref{tab:s1063parameters}, comparing our work here with parameters for S1063 determined in \citet{mathieu03}, \citet{leiner17}, and the SPOTS model comparison above. As the closest SPOTS model depends on the comparison photometry used (see Figure~\ref{fig:CMDs}), the stellar parameters for both models are given.

We further discuss systematic uncertainties that may impact these discrepancies in Section~\ref{sec:uncertainties}, but fully resolving the complexities of photometric and spectroscopic constraints of spot filling factors likely requires careful photometric and atmospheric modeling of a multi-temperature stellar surface, which is beyond the scope of this paper.

@@

The spot coverage fraction for S1063 determined in this work is consistent with the range of spot coverage seen on RS CVn of 30--40\% from measuring TiO band strength \citep{oneal96, oneal98, oneal04} and Doppler imaging \citep{hackman12}. This is further evidence that this SSG and likely other SSGs have high magnetic activity.

@@

\subsection{Effective temperature of spotted stars}
The conventional observational interpretation of $T_{\textrm{eff}}$ is difficult to apply to spotted stars, especially stars with variation in the spot covering fraction over time \citep{gullysantiago17}. Different methods of characterizing $T_{\textrm{eff}}$ complicate  comparisons between observations and theoretical models.

%Conventionally, a star is thought to have a single $T_{\textrm{eff}}$ that can be easily matched to  Conventionally, a spectrum can be interpreted as possessing a 1:1 mapping from its appearance to a $T_{\textrm{eff}}$-indexed template, which can be either model-based or observed.  We...

\citet{leiner17} fit the multi-band spectral energy distribution for S1063 with a single Castelli-Kurucz spectrum \citep{2003IAUS..210P.A20C} and find a best fit single-temperature $T_{\textrm{eff}}$ of 4500 K. With observational confirmation that S1063 is a spotted star, another interpretation is to calculate the surface-averaged $T_{\textrm{eff}}$:

\begin{equation}
T_{\textrm{eff}}^4 = f_{\textrm{s}} T_{\textrm{s}}^4 + (1 -f_{\textrm{s}}) T_{\textrm{amb}}^4 .
\end{equation}

where $f_{\textrm{s}}$ is the spot covering fraction, $T_{\textrm{s}}$ is the spot temperature, $T_{\textrm{amb}}$ is the ambient photosphere temperature. This surface-averaged temperature allows for comparisons to one-dimensional stellar evolution models motivated by the surface conditions of the star \citep{somers20}. With this method, at the time of the IGRINS observation the $T_{\textrm{eff}}$ of S1063 is $4900\pm100$ K. This temperature is inconsistent with the temperature determined via SED fitting in \citet{leiner17}, indicative of the discrepancies between the spectral and photometric constraints outlined in Section~\ref{sec:model_comparison}, although we note that the previous temperature of 4500 K is between the measured spot and ambient photosphere temperatures determined in this work.

Stars such as S1063, however, have observed surface-averaged $T_{\textrm{eff}}$ values that vary based on the spot covering fraction at the time of observation. We know from the variable light curve (Figure~\ref{fig:lightcurve}) that stars such as S1063 have differing spot covering fractions on opposite observable hemispheres of the star causing optical variability. For example, at the four-year light curve maximum S1063 had approximately 20\% spot covering fraction and a surface-averaged $T_{\textrm{eff}}$ (for the observed hemisphere of the star) of 5000 K. At the light curve minimum with a 45\% spot covering fraction the surface-averaged $T_{\textrm{eff}}$ for the observed hemisphere would only be 4700 K.

The true surface-averaged $T_{\textrm{eff}}$ includes the entire stellar surface. This should be constant over a thermal timescale, much longer than the four-year time frame in Figure~\ref{fig:lightcurve}, as is demonstrated by the constant mean flux level of 89.5\% compared to the maximum global flux. Observationally constraining the surface-averaged $T_{\textrm{eff}}$ requires either multi-epoch spectrum observations at both the light curve maximum and minimum, or a single observation at the mean flux level. The date of the IGRINS observation corresponds to a flux level of 91.2\% compared to the global maximum, suggesting that observed spot coverage is a reasonable proxy for the mean state of S1063. Therefore, we conclude that the total surface-averaged $T_{\textrm{eff}}$ of S1063 is consistent with the observed surface-averaged $T_{\textrm{eff}}$ of $4900\pm100$ K. We stress, however, that determining a single $T_{\textrm{eff}}$ value for spotted stars makes model comparisons exceedingly nuanced.

@@

The radius of S1063 is encoded into data in three ways: via surface gravity-sensitive spectral lines, in the SED and HR Diagram position, and embedded within the $v\sin{i}$ measurement. Observationally constraining the radius of spotted stars would help model comparisons, especially in light of the single $T_{\textrm{eff}}$ nuances elaborated above.

Our weak constraints on surface gravity from the IGRINS spectra cannot meaningfully inform the radius. Previously determined SED-based photometric results suggest a radius of $R_{\star} = 2.8-3.4 R_\odot$ (see Table~\ref{tab:s1063parameters}). We elaborate on constraining the radius from $v\sin{i}$ here.
The stellar radius $R_{\star}$ contributes to the magnitude of the equatorial velocity $v_{\mathrm{eq}}$ for a given rotation rate and projected stellar inclination angle $i$:
\begin{eqnarray}
  v \sin{i} = \frac{2\pi R_{\star}}{P_{\mathrm{rot}}} \sin{i} \\
  R_{\star} \sin{i} = \frac{v \sin{i} \cdot P_{\mathrm{rot}}}{2 \pi} \label{rsini}
\end{eqnarray}

Our \texttt{Starfish}/IGRINS-derived $v\sin{i}=10.0 \pm 1.0 \; \mathrm{km\,s^{-1}}$ and \texttt{celerite}/K2 derived period $P_{\mathrm{rot}}=23.5 \pm 0.1$ constrain $R\sin{i} = 4.6 \;R_\odot$, with a formal statistical uncertainty of $\pm 0.5 \;R_\odot$.  For stars viewed at an inclination other than perfectly edge-on, the typical inferred stellar radius $R_{\star}$ must exceed $4.6 R_{\odot}$. If we instead adopt the optically-determined $v\sin{i}=8\; \mathrm{km\,s^{-1}}$ from \citet{mathieu03} the radius must exceed 3.7 $R_{\odot}$. In either case, this kinematically-inferred radius is inconsistent with the radius range determined through photometric studies. A two-temperature SED of S1063 adopting the minimum $R\sin{i}$ values is far too bright compared to observed photometry, suggesting that the $R\sin{i}$ limits here are biased to larger values in some way.

One possible source of this discrepancy is the spotted surface somehow alters the assumptions behind Equation \ref{rsini}. Possible astrophysical uncertainties that could potentially bias the $v\sin{i}$ for heavily spotted sub-giants include the adopted limb-darkening prescription, micro- or macro-turbulence, and Zeeman broadening. Additionally, the observed period could be biased via differential rotation combined with starspots constrained to either high or low latitudes, starspot evolution timescales similar to the rotation period, and aliasing from regularly-spaced longitudinal surface symmetries. The magnitude and/or direction of these period biases are generally uncertain. We note that the other SSG in M67, S1113, also has an anomalously-large $R\sin{i}$ value determined from $v\sin{i}$ \citep{mathieu03}, suggesting that perhaps the surface conditions of SSGs cause spectral broadening such that the standard $v\sin{i}$ measurement is not truly a measure of the stellar rotation velocity. A campaign of intense doppler-imaging-style observations could act to disentangle some of these uncertainties in the future, though such a campaign would require extreme spectral resolution in order to resolve surface structures given the relatively low apparent $v\sin{i}$.

>>>>

Starspots have become a rich observational proxy for gauging the evolutionary state of stars that may be magnetically inflated without needing to estimate magnetic field strengths.  A full accounting of starspots requires careful consideration of geometrical degeneracies.  Monochromatic lightcurves can indicate spotted stars \citep{2014ApJS..211...24M} but only provide a lower limit on the differential spot coverage between the most and least spotted hemispheres. Longitudinally symmetric starspot geometries evade detection in monochromatic lightcurves alone \citep{2019AJ....157...64L}. Disk-integrated covering fractions can be obtained from TiO band observations \citep{oneal96,fang2016,2019AJ....158..101M}, but measuring the spot configuration typically requires Doppler imaging or interferometry studies restricted to only the brightest and most nearby sources \citep{roettenbacher16}.  Testing the next era of stellar activity models requires precision methodologies to measure spot-covering fractions of a larger number of targets, including SSGs.

Starspots emit a spectrum of their own at a lower temperature and with distinct absorption features compared to the ambient photosphere. The observed spectrum is therefore a composite of the spot and photosphere spectra, which can be deconvolved.  \citet{gullysantiago17} extend the \texttt{Starfish} \citep{czekala15} spectral inference framework to support composite spectra and two-temperature probabilistic spectral decomposition. This deconvolution process constrains the spot-covering fraction as well as the spot and photosphere temperatures. The resulting disk-average spot properties anchor a simultaneous lightcurve flux to the starspot coverage fraction. This anchored value serves as a benchmark for interpreting long-term (multi-year) flux variability \citep{neff95}. This strategy only requires high-resolution near-IR echelle spectroscopy and photometric monitoring, making it less demanding on telescope resources than either Doppler Imaging or interferometry. The technique is amenable to sources with narrow spectral lines ($v\sin{i}<10$ km/s) and stars with any stellar inclination, a distinct advantage over Doppler imaging.  Altogether these advantages dramatically increase the number of sources for which we can observationally constrain spot-covering fractions.

In this paper we demonstrate the power of this methodology by focusing on a single SSG system, S1063 \citep{1977A&AS...27...89S} in the open cluster M67. This system is a prototypical SSG, with a single-lined spectroscopic orbital period of 18.4 days \citep{geller17}, an X-ray luminosity of $1.3\times10^{31}$ erg s$^{-1}$ \citep{vandenberg99}, and a variable lightcurve in \textit{K2} and ASAS-SN (Section~\ref{sec:K2lightcurve}). \emph{Gaia} DR2 astrometry \citep{2016A&A...595A...1G, 2018A&A...616A...1G} indicates a parallax ($1.17\pm0.025 \;$mas) and proper motion for S1063 (\emph{Gaia} DR2 604921030968952832) consistent with other M67 members  approaching 100\% membership probability \citep{2018ApJ...869....9G}. We present our observational data products in Section~\ref{sec:observations} and subsequent analysis in Section~\ref{sec:analysis}. Our results are given in Section~\ref{sec:results} with Discussion in Section~\ref{sec:discussion} including additional discussion of the possible systematic uncertainties that may bias our results. Finally, our conclusions are outlined in Section~\ref{sec:conclusions}.

@@

To account for sensitivity changes, we calibrate the system-integrated throughput for \emph{Kepler} detector channels including the M67 field for Campaigns 5, 16, and 18. We measure aperture photometry for approximately 2000 isolated reference stars from the Full Frame Images, keeping only those stars that were observed in all three campaigns. Compared to Campaign 5, the reference stars have a median flux of $93.9\pm4.2\%$ in Campaign 16 and a median flux of $98.2\pm2.8\%$ in Campaign 18. Campaigns 5 and 18 were observed on the same detector channel while a different detector channel was used for Campaign 16, therefore the C16 offset is not a significant measure of detector degradation across campaigns. We use a Lomb-Scargle periodogram on each campaign independently and recover the same dominant period of 23.5 days in all three campaigns. The phase-folded lightcurves and periodograms are shown in Figure~\ref{fig:periodogram}.

\begin{figure*}[ht]
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=3in]{figures/AllCampaigns_Periodogram.pdf}   & \includegraphics[width=3in]{figures/AllCampaigns_Phased_Lightcurve.pdf}  \\
    \end{tabular}
    \caption{On the right, a Lomb-Scargle Periodogram of the lightcurve for each \emph{K2} campaign and on the left, the corresponding phased \emph{K2} lightcurves. The average period associated with maximum power across all three campaigns is P = 23.5 days.  }
    \label{fig:periodogram}
\end{figure*}

\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.95\textwidth]{figures/2020_K2_ASASSN_lcurve_2panel.pdf}
\caption{Four year lightcurve for S1063, normalized to a maximum flux of 1.00 as denoted by the thick gray dashed line. The mean flux of 0.895 is shown with the gray dotted line.  \emph{K2} Campaigns 5, 16, and 18 (densely-sampled red points) and ASAS-SN $V-$band (coarsely-sampled gray points) show approximately $3-17\%$ peak-to-valley photometric variations indicative of secularly evolving surface coverage of starspots.  The orange line shows a damped, driven harmonic oscillator \texttt{celerit\`e} model \citep{2017AJ....154..220F} trained on the \emph{K2} and ASAS-SN data, and then applied to the noisy ASAS-SN points. A standard-error uncertainty band is shown in shaded orange. The vertical blue bar in mid-2015 indicates the epoch of IGRINS data acquisition, which shows an approximate $8\%$ flux deficit compared to the global maximum flux in early 2017.}
\label{fig:lightcurve}
\end{figure*}

@@

%INTRO TEXT:
%Starspots emit a spectrum of their own at a lower temperature and with distinct absorption features compared to the ambient photosphere. The observed spectrum is a composite of the spot and photosphere spectra and can therefore be deconvolved. This deconvolution constrains the spot-covering fraction as well as the spot and photosphere temperatures. This is possible by performing a two-temperature probabilistic spectral decomposition using the spectral inference framework \texttt{Starfish} \citep{czekala15}, recently extended to support composite spectra \citep{gullysantiago17}. The resulting spot characterization can be used to anchor lightcurve information, providing insight into the overall stellar activity level and constraining the long-term spot coverage \citep{neff95}. This strategy only requires high-resolution echelle spectroscopy and photometric monitoring, dramatically increasing the number of sources for which we can observationally constrain spot-covering fractions.

%Constraining starspot covering fraction for stars such as S1063 requires a careful combination of lightcurve interpretation and spectral analysis using spectral decomposition tools such as \texttt{Starfish} \citep{czekala15} to benchmark epochs of starspot coverage in long-term photometric monitoring.
Characterizing the long-term starspot coverage of S1063 requires connecting coarse ground-based and precision space-based photometry---ASAS-SN and \emph{K2}---alongside precision IGRINS spectroscopy.  Here we lay out the calibration of these datasets to accommodate their wide range in data quality.

@@

For a first-order interpretation of lightcurve amplitudes, we posit that the lightcurve is starspot-dominated as opposed to facula-dominated.  Under this assumption, minimum light corresponds to the largest blockage of ambient flux and therefore the largest starspot coverage on the instantaneous hemisphere projected towards the observer \citep{basri18}.  Maximum light corresponds to the moments with the least---\emph{but not necessarily zero}---starspot coverage.

@@

This 13\% spot coverage is a minimum value based on an assumption that the global maximum lightcurve flux results from an un-spotted star. Our spectral inference results indicate the spot coverage at the time of the IGRINS observation was 32\%.  Therefore the spot coverage of S1063 was approximately 20\% at maximum light over this 4-year period, and not zero as a first-order interpretation of the lightcurve would imply.

The presence of starspots at the lightcurve maximum acts as a starspot baseline for interpreting the lightcurve. Over the four years shown in Figure~\ref{fig:lightcurve}, the average flux variation is approximately 5\%, corresponding to an average spot covering fraction variation from each observed hemisphere of $\pm$8\%. The lightcurve exhibits a global flux minimum 10\% lower than the flux at the time of the IGRINS observation, requiring a spot filling factor close to 45\% on the projected hemisphere at that time.

We therefore determine that over the time period from 2014--2018 S1063 has possessed a range of 20--45\% coverage fraction of spots, with an average close to 30\%. The IGRINS observation occurring near the mean global flux level fortuitously allows a robust characterization of the average state of S1063. These numbers possess formal uncertainties in the few percent range, and additional systematic uncertainties from our reliance on spectral models.  But broadly speaking, our result is firm under our assumptions --- S1063 has a large persistent starspot population that ebbs and flows in its longitudinal symmetry, resulting in variation in the peak-to-valley modulation, but routinely possessing an approximately 30\% spot covering fraction.

@@

We believe that some of the discrepancy between our results and those predicted by the SPOTS models may be due to limitations of the SPOTS models as diagnostic tools for single sources. However, observational biases may also impact the filling factors determined through spectroscopic and photometric methods. Our near-IR spectral decomposition technique is most sensitive to detecting a ``sweet-spot'' in starspot contrast: cool enough to exhibit a spectrum that is qualitatively distinct from the ambient photosphere, but not so cool as to possess vanishing specific intensity.  So for example, our technique might readily perceive the penumbra of the starspot, while ignoring the much cooler and darker umbra flux contribution \citep{1981ApJ...250..327V}.  This bias would lead our technique to systematically underestimate the total starspot coverage, rendering our claimed filling factor estimate as a lower limit. Spectral decomposition techniques are more sensitive to the warmer spot components and may therefore be constraining the prenumbra filling factor, which would be larger than the umbra filling factor. However, we do not know if the spot morphology of SSGs follows this assumption. If instead SSGs have a smaller ratio of prenumbra to umbra this spectral technique would under-sample the total spot coverage on the stellar surface.

Other second-order effects could be at play.  Spots could hypothetically exhibit a perceived wavelength dependence of their boundaries that would lead to apparent discrepancies in spot filling factors across photometric and spectroscopic determinations at different wavebands.  Such effects are not well-understood for stars other than the Sun, where the exact boundaries of active regions can be examined in different filters.  We note that photometry collected at different epochs can also degrade the precision of SED-based photometric filling factors. Given our lightcurve analysis of S1063 we expect SED-based filling factor uncertainties could be as large as $\sim$20\% from unluckily sampling peaks and troughs of the lightcurve.

A summary of the surface and stellar parameters for S1063 are given in Table~\ref{tab:s1063parameters}, comparing our work here with parameters for S1063 determined in \citet{mathieu03}, \citet{leiner17}, and the SPOTS model comparison above. As the closest SPOTS model depends on the comparison photometry used (see Figure~\ref{fig:CMDs}), the stellar parameters for both models are given.  Fully resolving the complexities of photometric and spectroscopic constraints of spot filling factors likely requires careful photometric and atmospheric modeling of a multi-temperature stellar surface, which is beyond the scope of this paper.

@@

The spot coverage fraction for S1063 determined in this work is consistent with the range of spot coverage seen on RS CVn of 30--40\% from measuring TiO band strength \citep{oneal96, oneal98, oneal04} and Doppler imaging \citep{hackman12}. The collective evidence supports the interpretation that S1063 and by extension other SSGs have magnetic activity so severe as to alter their positions in the HR diagram.

@@

\subsection{Effective temperature of spotted stars}

%% Is the first sentence too obvious?
Effective temperature relates the radius and luminosity of a star through the Stefan-Boltzmann Law.  The radius of S1063 remains constant over thermal timescales much longer than the span of available data.  We assume the long-time averaged luminosity also remains constant as starspots come and go on the surface, but have a near-constant time-averaged occupation fraction.  A hint of this quasi-static luminosity state can be conceptualized in Figure \ref{fig:lightcurve} with a relatively constant mean flux level of 89.5\% compared to the maximum global flux.  In these limits, there exists a well-defined $T_{\textrm{eff}}$ that conventionally relates the long-time-averaged luminosity and radius:
\begin{equation}
T_{\textrm{eff}}^4 = f_{\textrm{s}} T_{\textrm{s}}^4 + (1 -f_{\textrm{s}}) T_{\textrm{amb}}^4 . \label{defineTeff}
\end{equation}
where $f_{\textrm{s}}$ is the long-time-averaged spot covering fraction, $T_{\textrm{s}}$ is the long-time-averaged spot temperature, $T_{\textrm{amb}}$ is the long-time-averaged ambient photosphere temperature.  Two factors make it challenging to measure and interpret effective temperature of spotted stars in practice. First, the number of surface components may be misassigned.  For example \citet{leiner17} fit the multi-band spectral energy distribution for S1063 with a single Castelli-Kurucz spectrum \citep{2003IAUS..210P.A20C} and find a best fit single-temperature $T_{\textrm{eff}}$ of 4500 K.  Here we show that two temperatures is better justified.  It may be that 3 components is even better justified, requiring a further adjustment to Equation \ref{defineTeff}.   Second, instantaneously derived effective temperatures may be acquired at a time that is not representative of the steady-state.

Here we attempt to compare the surface-averaged temperature to one-dimensional stellar evolution models motivated by the surface conditions of the star \citep{somers20}. At the time of the IGRINS observation the $T_{\textrm{eff}}$ of S1063 derived with Equation \ref{defineTeff} is $4900\pm100$ K. This temperature is inconsistent with the 4500 K temperature determined via single-temperature SED fitting in \citet{leiner17}, indicative of the discrepancies between the spectral and photometric constraints outlined in Section~\ref{sec:model_comparison}, although we note that the previous temperature of 4500 K lands between the measured spot and ambient photosphere temperatures determined in this work.

Stars such as S1063 can have instantaneously-derived $T_{\textrm{eff}}$ values that vary based on the spot covering fraction at the time of observation. We know from the variable lightcurve (Figure~\ref{fig:lightcurve}) that stars such as S1063 have differing spot covering fractions on opposite observable hemispheres of the star causing optical variability. For example, at the four-year lightcurve maximum S1063 had approximately 20\% spot covering fraction and an instantaneously-derived surface-averaged $T_{\textrm{eff}}$ (for the observed hemisphere of the star) of 5000 K. At the lightcurve minimum with a 45\% spot covering fraction the surface-averaged $T_{\textrm{eff}}$ for the observed hemisphere would only be 4700 K.

The true surface-averaged $T_{\textrm{eff}}$ includes the entire stellar surface. Observationally constraining the surface-averaged $T_{\textrm{eff}}$ requires either multi-epoch spectrum observations at both the lightcurve maximum and minimum, or a single observation at the mean flux level. The date of the IGRINS observation corresponds to a flux level of 91.2\% compared to the global maximum, suggesting that observed spot coverage is a close proxy for the mean state of S1063 within the uncertainties. Therefore, we conclude that the total surface-averaged $T_{\textrm{eff}}$ of S1063 is consistent with the observed surface-averaged $T_{\textrm{eff}}$ of $4900\pm100$ K. We stress, however, that determining a single $T_{\textrm{eff}}$ value for spotted stars makes model comparisons exceedingly nuanced.  Stars like S1063 may undergo stellar cycles with possibly massive changes to spot coverage over timescales of hundreds of years, much longer than our observational windows.

\subsection{Flares}

A star as allegedly active and spotted as S1063 should exhibit other metrics associated with activity beyond the already-mentioned hallmarks such as H$\alpha$, x-ray emission, and starspot coverage.  We should see flares.  A closer re-examination of Figures \ref{fig:periodogram} and \ref{fig:lightcurve} reveals conspicuous stellar flares in all three of the campaigns of \emph{K2}.  The mere existence of these flares conveys supporting evidence for the magnetic interpretation for SSGs.  The detection of flares also hints at a promising prediction for the analysis of known and undiscovered SSGs: \emph{TESS} lightcurves of sub-subgiants should exhibit a comparable frequency and amplitude of stellar flares as those seen in S1063.  We may also expect a difference in the frequency of flares between SSGs and RS CVn stars if these two populations possess inherently different degrees of magnetic activity.  Such a comparison will benefit from the much larger sample sizes of SSGs afforded by the all-sky \emph{TESS} mission in concert with \emph{Gaia}.

@@

The radius of S1063 is encoded into data in three ways: via surface gravity-sensitive spectral lines, in the SED and HR Diagram position, and embedded within the $v\sin{i}$ measurement. Measuring the radius of spotted stars would help constrain magnetic inflation models, especially in light of the single $T_{\textrm{eff}}$ nuances elaborated above.

Our weak constraints on surface gravity from the IGRINS spectra cannot meaningfully inform the radius without mass. Previously determined SED-based photometric results suggest a radius of $R_{\star} = 2.8-3.4 R_\odot$ (see Table~\ref{tab:s1063parameters}). We elaborate on constraining the radius from $v\sin{i}$ here.
The stellar radius $R_{\star}$ contributes to the magnitude of the equatorial velocity $v_{\mathrm{eq}}$ for a given rotation rate and projected stellar inclination angle $i$:
\begin{eqnarray}
  R_{\star} \sin{i} = \frac{v \sin{i} \cdot P_{\mathrm{rot}}}{2 \pi} \label{rsini}
\end{eqnarray}

Our \texttt{Starfish}/IGRINS-derived $v\sin{i}=10.0 \pm 1.0 \; \mathrm{km\,s^{-1}}$ and \texttt{celerite}/K2 derived period $P_{\mathrm{rot}}=23.5 \pm 0.1$ constrain $R\sin{i} = 4.6 \;R_\odot$, with a formal statistical uncertainty of $\pm 0.5 \;R_\odot$.  For stars viewed at an inclination other than perfectly edge-on, the typical inferred stellar radius $R_{\star}$ must exceed $4.6 R_{\odot}$. If we instead adopt the optically-determined $v\sin{i}=8\; \mathrm{km\,s^{-1}}$ from \citet{mathieu03} the radius must exceed 3.7 $R_{\odot}$. In either case, this kinematically-inferred radius is larger than the radius range determined through photometric studies. A two-temperature SED of S1063 adopting the minimum $R\sin{i}$ values is far too bright compared to observed photometry, suggesting our $R\sin{i}$ limits may be biased to larger values in some way.

One possible source of this radius discrepancy is the spotted surface somehow alters the assumptions behind Equation \ref{rsini}. Possible astrophysical uncertainties that could potentially bias the $v\sin{i}$ for heavily spotted sub-giants include the adopted limb-darkening prescription, micro- or macro-turbulence, and Zeeman broadening. Additionally, the observed period could be biased via differential rotation combined with starspots constrained to either high or low latitudes, starspot evolution timescales similar to the rotation period, and aliasing from regularly-spaced longitudinal surface symmetries. The magnitude and/or direction of these period biases are generally uncertain. We note that the other SSG in M67, S1113, also has an anomalously-large $R\sin{i}$ value determined from $v\sin{i}$ \citep{mathieu03}, suggesting that perhaps the surface conditions of SSGs cause spectral broadening such that the standard $v\sin{i}$ measurement is not truly a measure of the stellar rotation velocity. A campaign of intense Doppler-imaging-style observations could act to disentangle some of these uncertainties in the future, though such a campaign would require extreme spectral resolution in order to resolve surface structures given the relatively low apparent $v\sin{i}$.
